---
title: "Nonstationary SST-salmon relationships"
author: "Mike Litzow"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
library(MuMIn)
library(dplyr)
library(ggplot2)
library(nlme)
library(lmtest)
options(max.print = 99999)
```

#Approach
Here I used mixed-effects models to test for changing SST effects by era. I fit separate models to each species, combining data over all three regions. I'm using the approach outlined by Zuur et al. 2009:

First, fit model with full fixed effects structure, using REML estimation, and select best random effects structure using AICc.
Second, refit with ML and compare nested fixed effects models to evaluate support for different fixed terms. This comparison uses the likelihood ratio test.
Third, refit selected model with REML to get unbiased coefficient estimates.

The full fixed structure that I'm using is:
ln(R/S) ~ intercept + stock:spawners + sst:region + sst:region:era 

Note that this model invokes a nonstationary sst effect for each region separately. I then test the nonstationary hypothesis by comparing with this reduced model:
ln(R/S) ~ intercept + stock:spawners + sst:region

The reduced model invokes a stationary sst effect for each region.

I'm using lme with autocorrelated residuals throughout. The support for autocorrelated residuals is evaluated with another round of likelihood ratio testing.

Region = EBS/GOA/South, sst = sst at ocean entry, era = entry year before/after 1988/89.

Also...comparing the three different possible groupings of SST months for use in analysis, generally from broadest (SST1) to most restricted (SST2).

#Pink salmon
Begin by fitting the random effects. I'll compare a random intercept model, and random structure invoking each of the two interaction terms (sst:region and sst:region:era), both with and without intercepts.

```{r}
run.dat <- read.csv("coastwide salmon data with different sst groupings.csv", row.names = 1)

formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region + sst1:region:era)

# fixed effects model
p1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept for stocks
p2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# full interaction with intercept
p3 <- lme(formula.full, random = ~1 + sst1:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
p4 <- lme(formula.full, random = ~1 + sst1:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
p5 <- lme(formula.full, random = ~ -1 + sst1:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
p6 <- lme(formula.full, random = ~-1 + sst1:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Pink"))

AIC <- AICc(p1, p2, p3, p4, p5, p6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Random intercepts random structure is the best model by far. Look at the residuals:

```{r, echo=F}
qqnorm(p2, ~ranef(., level=1))
```

Pretty good? 

Now the fixed effects structure.

```{r}
# full model is p2 from above - but changing to ML estimation
pfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region)
p1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

anova(pfull, p1)
```

The sst1:region:era term is strongly supported. I'll compare with the model without autocorrelated residuals.

```{r}
# alternate model without AR(1) residuals
p2 <- lme(formula.full, random = ~1 | stock,
          method = "ML", data = filter(run.dat, species=="Pink"))

anova(pfull, p2)
```

So there isn't any big difference in the AIC scores, but p2 is better, though there is no support in the LRT for retaining the autocorrelated structure. For uniformity with the other spp., where AR(1) residuals are well supported, I'll retain that structure here. Also, will save as p.sst1 for model comparison.

```{r}
p.sst1 <- pfull
```

Here is the t-table.

```{r, echo=F}
pfull <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))
summary(pfull)$tTable[37:42,]
```

Significant sst1:era effects for GOA and South.

And now the plot.

```{r, echo=F}

pplot <- as.data.frame(intervals(pfull, which="fixed")$fixed[c(37:42),])

pplot$term <- rownames(pplot)

pplot$term <- reorder(pplot$term, pplot$est.)

ggplot(pplot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Pink salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))
```

#Sockeye salmon

Same routine - begin with random effects. 

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region + sst1:region:era)

# fixed effects model
s1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# random intercept for stocks
s2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# full interaction with intercept
s3 <- lme(formula.full, random = ~1 + sst1:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
s4 <- lme(formula.full, random = ~1 + sst1:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
s5 <- lme(formula.full, random = ~ -1 + sst1:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
s6 <- lme(formula.full, random = ~-1 + sst1:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Sockeye"))

AIC <- AICc(s1, s2, s3, s4, s5, s6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Again, model 2 massively better. Here are the random effect residuals.

```{r, echo=F}
qqnorm(s2, ~ranef(., level=1))
```

Now the hypothesis test via comparing fixed effects.

```{r}
# full model from above - but changing to ML estimation
sfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region)
s1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

anova(sfull, s1)
```

The sst1:region:era term is supported. I'll compare with the model without autocorrelated residuals.

```{r}

s2 <- lme(formula.full, random = ~1 | stock,
          method = "ML", data = filter(run.dat, species=="Sockeye"))

anova(sfull, s2)
```

In this case the AR(1) errors are far better. Save ML version for comparison below.

```{r}
s.sst1 <- sfull
```


Here is the t-table.

```{r, echo=F}
sfull <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))
summary(sfull)$tTable[32:37,]
```

And the plot.

```{r, echo=F}

splot <- as.data.frame(intervals(sfull, which="fixed")$fixed[c(32:37),])

splot$term <- rownames(splot)

splot$term <- reorder(splot$term, splot$est.)

ggplot(splot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Sockeye salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))
```

#Chum salmon

Random effects.

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region + sst1:region:era)

# fixed effects model
c1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# random intercept for stocks
c2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# full interaction with intercept - DOESN'T FIT!
# c3 <- lme(formula.full, random = ~1 + sst1:region:era | stock,
#           method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
#           control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
c4 <- lme(formula.full, random = ~1 + sst1:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
c5 <- lme(formula.full, random = ~ -1 + sst1:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
c6 <- lme(formula.full, random = ~-1 + sst1:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Chum"))

AIC <- AICc(c1, c2, c4, c5, c6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Model 2 again.

```{r, echo=F}
qqnorm(c2, ~ranef(., level=1))
```

And fixed effects.

```{r}
# full model - changing to ML estimation
cfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region)
c1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c1)
```

Nonstationary model supported. Compare different error structures.

```{r}

cfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

c2 <- lme(formula.full, random = ~1 | stock,
          method = "ML", data = filter(run.dat, species=="Chum"))

anova(cfull, c2)
```

Again, the AR(1) errors are far better!

Save the ML version.
```{r}
c.sst1 <- cfull
```


Here is the t-table.

```{r, echo=F}
cfull <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))
summary(cfull)$tTable[23:28,]
```

```{r}
coef(cfull)
```


And plot.

```{r, echo=F}


cplot <- as.data.frame(intervals(cfull, which="fixed")$fixed[c(23:28),])

cplot$term <- rownames(cplot)

cplot$term <- reorder(cplot$term, cplot$est.)

ggplot(cplot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Chum salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))
```

#Now repeat with SST2.

#Pink salmon
Begin by fitting the random effects. I'll compare a random intercept model, and random structure invoking each of the two interaction terms (sst:region and sst:region:era), both with and without intercepts.

```{r}

formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region + sst2:region:era)

# fixed effects model
p1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept for stocks
p2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# full interaction with intercept
p3 <- lme(formula.full, random = ~1 + sst2:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
p4 <- lme(formula.full, random = ~1 + sst2:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
p5 <- lme(formula.full, random = ~ -1 + sst2:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
p6 <- lme(formula.full, random = ~-1 + sst2:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Pink"))

AIC <- AICc(p1, p2, p3, p4, p5, p6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Random intercepts random structure is the best model by far. Look at the residuals:

```{r, echo=F}
qqnorm(p2, ~ranef(., level=1))
```

Now the fixed effects structure.

```{r}
# full model is p2 from above - but changing to ML estimation
pfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region)
p1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

anova(pfull, p1)
```

The sst2:region:era term is also strongly supported. Save best model as p2 for comparison.

```{r, echo=F}
p.sst2 <- pfull
```

#Sockeye salmon

Same routine - begin with random effects. 

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region + sst2:region:era)

# fixed effects model
s1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# random intercept for stocks
s2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# full interaction with intercept
s3 <- lme(formula.full, random = ~1 + sst2:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
s4 <- lme(formula.full, random = ~1 + sst2:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
s5 <- lme(formula.full, random = ~ -1 + sst2:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
s6 <- lme(formula.full, random = ~-1 + sst2:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Sockeye"))

AIC <- AICc(s1, s2, s3, s4, s5, s6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Same random structure is best.

```{r, echo=F}
qqnorm(s2, ~ranef(., level=1))
```

Now the hypothesis test via comparing fixed effects.

```{r}
# full model from above - but changing to ML estimation
sfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region)
s1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

anova(sfull, s1)
```

Nonstationary model supported, but not as strongly. Save for model comparison.

```{r}
s.sst2 <- sfull
```

#Chum salmon

Random effects.

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region + sst2:region:era)

# fixed effects model
c1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# random intercept for stocks
c2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# full interaction with intercept - DOESN'T FIT!
# c3 <- lme(formula.full, random = ~1 + sst2:region:era | stock,
#           method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
#           control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
c4 <- lme(formula.full, random = ~1 + sst2:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
c5 <- lme(formula.full, random = ~ -1 + sst2:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
c6 <- lme(formula.full, random = ~-1 + sst2:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Chum"))

AIC <- AICc(c1, c2, c4, c5, c6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Model 2 again.

```{r, echo=F}
qqnorm(c2, ~ranef(., level=1))
```

And fixed effects.

```{r}
# full model - changing to ML estimation
cfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst2:region)
c1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c1)
```

Nonstationary model supported. Save.

```{r}
c.sst2 <- cfull
```

#Finally, SST3.

#Pink salmon
Begin by fitting the random effects. I'll compare a random intercept model, and random structure invoking each of the two interaction terms (sst:region and sst:region:era), both with and without intercepts.

```{r}

formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region + sst3:region:era)

# fixed effects model
p1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept for stocks
p2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# full interaction with intercept
p3 <- lme(formula.full, random = ~1 + sst3:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
p4 <- lme(formula.full, random = ~1 + sst3:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
p5 <- lme(formula.full, random = ~ -1 + sst3:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
p6 <- lme(formula.full, random = ~-1 + sst3:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Pink"))

AIC <- AICc(p1, p2, p3, p4, p5, p6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Random intercepts random structure is the best model by far. Look at the residuals:

```{r, echo=F}
qqnorm(p2, ~ranef(., level=1))
```


Now the fixed effects structure.

```{r}
# full model is p2 from above - but changing to ML estimation
pfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region)
p1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

anova(pfull, p1)
```

The sst3:region:era term is also strongly supported. Save best model as p3 for comparison.

```{r, echo=F}
p.sst3 <- pfull
```

#Sockeye salmon

Same routine - begin with random effects. 

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region + sst3:region:era)

# fixed effects model
s1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# random intercept for stocks
s2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# full interaction with intercept
s3 <- lme(formula.full, random = ~1 + sst3:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept
s4 <- lme(formula.full, random = ~1 + sst3:region | stock,
         method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
         control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
s5 <- lme(formula.full, random = ~ -1 + sst3:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
s6 <- lme(formula.full, random = ~-1 + sst3:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Sockeye"))

AIC <- AICc(s1, s2, s3, s4, s5, s6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Same random structure is best.

```{r, echo=F}
qqnorm(s2, ~ranef(., level=1))
```

Now the hypothesis test via comparing fixed effects.

```{r}
# full model from above - but changing to ML estimation
sfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region)
s1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

anova(sfull, s1)
```

Nonstationary model supported, but more weakly. Save for model comparison.

```{r}
s.sst3 <- sfull
```

#Chum salmon

Random effects.

```{r}
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region + sst3:region:era)

# fixed effects model
c1 <- gls(formula.full, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# random intercept for stocks
c2 <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# full interaction with intercept - DOESN'T FIT!
# c3 <- lme(formula.full, random = ~1 + sst3:region:era | stock,
#           method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
#           control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction with intercept - DOESN'T FIT!
# c4 <- lme(formula.full, random = ~1 + sst3:region | stock,
#          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
#          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# full interaction, no intercept
c5 <- lme(formula.full, random = ~ -1 + sst3:region:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# reduced interaction, no intercept
c6 <- lme(formula.full, random = ~-1 + sst3:region | stock,
          method = "REML", correlation=corAR1(form= ~1 | stock), data = filter(run.dat, species=="Chum"))

AIC <- AICc(c1, c2, c4, c5, c6) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Model 2 again.

```{r, echo=F}
qqnorm(c2, ~ranef(., level=1))
```

And fixed effects.

```{r}
# full model - changing to ML estimation
cfull <- lme(formula.full, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

formula.reduced <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region)
c1 <- lme(formula.reduced, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c1)
```

Nonstationary model supported. Save.

```{r}
c.sst3 <- cfull
```

#Model comparison for support of different possible SST groupings.

Use AICc selection to compare SST1, SST2, SST3 for each species.

First, pink.
```{r}
pink.aic <- AICc(p.sst1, p.sst2, p.sst3) %>%
  mutate(dAICc=AICc-min(AICc), model=c("p.sst1", "p.sst2", "p.sst3")) %>%
  arrange(dAICc) %>%
  print()

sock.aic <- AICc(s.sst1, s.sst2, s.sst3) %>%
  mutate(dAICc=AICc-min(AICc), model=c("s.sst1", "s.sst2", "s.sst3")) %>%
  arrange(dAICc) %>%
  print()

chum.aic <- AICc(c.sst1, c.sst2, c.sst3) %>%
  mutate(dAICc=AICc-min(AICc), model=c("c.sst1", "c.sst2", "c.sst3")) %>%
  arrange(dAICc) %>%
  print()
```

Finally, combine into a plot for the paper, using the best-supported sst range for each. Also, get the t tables for each best-supported model.

```{r}

# fit the best model for each species using REML
# sst3 is the best for pinks
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst3:region + sst3:region:era)
pbest <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# sst1 is the best for chums and sockeye
formula.full <- 
  as.formula(log(recruits/spawners) ~ 1 + stock:spawners + sst1:region + sst1:region:era)
sbest <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))
cbest <- lme(formula.full, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

summary(pbest)$tTable[37:42,]

summary(sbest)$tTable[32:37,]

summary(cbest)$tTable[23:28,]
```

And plot.

```{r}
plot <- data.frame(species=rep(c("Pink", "Sockeye", "Chum"), each=6), region=rep(c("EBS", "GOA", "South"), each=2),
                   era=c("Before 1988/89", "After 1988/89"), LCI=NA, effect=NA, UCI=NA)

plot[1,4:6] <- intervals(pbest, which="fixed")$fixed[37,] # pink EBS early 
plot[2,4:6] <- intervals(pbest, which="fixed")$fixed[37,2]+ 
  intervals(pbest, which="fixed")$fixed[40,] # pink EBS late

plot[3,4:6] <- intervals(pbest, which="fixed")$fixed[38,] # pink GOA early
plot[4,4:6] <- intervals(pbest, which="fixed")$fixed[38,2]+ 
  intervals(pbest, which="fixed")$fixed[41,] # pink GOA late

plot[5,4:6] <- intervals(pbest, which="fixed")$fixed[39,] # pink South early
plot[6,4:6] <- intervals(pbest, which="fixed")$fixed[39,2]+ 
  intervals(pbest, which="fixed")$fixed[42,] # pink South late

plot[7,4:6] <- intervals(sbest, which="fixed")$fixed[32,] # sockeye EBS early
plot[8,4:6] <- intervals(sbest, which="fixed")$fixed[32,2] + 
  intervals(sbest, which="fixed")$fixed[35,] # sockeye EBS late

plot[9,4:6] <- intervals(sbest, which="fixed")$fixed[33,] # sockeye GOA early
plot[10,4:6] <- intervals(sbest, which="fixed")$fixed[33,2] + 
  intervals(sbest, which="fixed")$fixed[36,] # sockeye GOA late

plot[11,4:6] <- intervals(sbest, which="fixed")$fixed[34,] # sockeye South early
plot[12,4:6] <- intervals(sbest, which="fixed")$fixed[34,2] + 
  intervals(sbest, which="fixed")$fixed[37,] # sockeye South late

plot[13,4:6] <- intervals(cbest, which="fixed")$fixed[23,] # chum EBS early
plot[14,4:6] <- intervals(cbest, which="fixed")$fixed[23,2] + 
  intervals(cbest, which="fixed")$fixed[26,] # chum EBS late 

plot[15,4:6] <- intervals(cbest, which="fixed")$fixed[24,] # chum GOA early
plot[16,4:6] <- intervals(cbest, which="fixed")$fixed[24,2] + 
  intervals(cbest, which="fixed")$fixed[27,] # chum GOA late

plot[17,4:6] <- intervals(cbest, which="fixed")$fixed[25,] # chum South early
plot[18,4:6] <- intervals(cbest, which="fixed")$fixed[25,2] + 
  intervals(cbest, which="fixed")$fixed[28,] # chum South late

plot$era <- reorder(plot$era, -order(plot$era))

plot$species <- reorder(plot$species, rep(c(1,2,3), each=6))

dodge <- position_dodge(width=0.9)

# load the color-blind palette
cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

## make a df for asterisks
label.df <- data.frame(species=rep(c("Pink", "Sockeye", "Chum"), each=3), 
                       region=c("EBS", "GOA", "South"), effect=c(1,1.35,1.8,0.5,0.5,-1.7,1,1,1.18), 
                       era=as.factor(rep(1,9)), 
                       label=c("", "***", "", "", "", "**", "", "", "**"))

ggplot(plot, aes(region, effect, fill=era)) + geom_hline(yintercept = 0, color="dark grey") + 
  geom_bar(position=dodge, stat="identity") + 
  geom_errorbar(aes(ymax=UCI, ymin=LCI), position=dodge, width=0.5) + xlab("") + 
  ylab("SST coefficient") + 
  facet_wrap(~species, scales="free_y") + 
  theme(legend.position = c(0.18, 0.85), legend.title = element_blank()) +
scale_fill_manual(values=cb[c(2,6,2)], breaks=c("Before 1988/89", "After 1988/89", "")) + 
  geom_text(data = label.df, label = label.df$label, size=6)

```

```{r, echo=F}
png("mixed effects results.png", 6,4, units="in", res=300)
ggplot(plot, aes(region, effect, fill=era)) + 
  theme_linedraw() +
  geom_bar(position=dodge, stat="identity") + geom_hline(yintercept = 0, color="black", lwd=0.3) + 
  geom_errorbar(aes(ymax=UCI, ymin=LCI), position=dodge, width=0.3, size=0.4) +
  xlab("") + ylab("SST coefficient") + 
  facet_wrap(~species, scales="free_y") + 
  theme(legend.position = c(0.18, 0.85), legend.title = element_blank()) +
scale_fill_manual(values=cb[c(2,6,2)], breaks=c("Before 1988/89", "After 1988/89", "")) + geom_text(data = label.df, label = label.df$label, size=6)

dev.off()
```

Make a quick plot of temperature distributions by era!

Using the best SST covariate for each species.

```{r}

plot.dat <- run.dat %>%
  select(species, region, sst1, sst3, era)
  
plot.dat$era <- ifelse(plot.dat$era==1, "Pre", "Post")
plot.dat$region.era <- paste(plot.dat$region, plot.dat$era, sep=" ")

plot.dat$sst.best <- ifelse(plot.dat$species=="Pink", plot.dat$sst1, plot.dat$sst3)

png("sst histograms.png", 6,6, units="in", res=300)
bw=20
ggplot(plot.dat, aes(sst.best)) + geom_histogram(bins = bw, fill="dark grey", color="black") + 
  facet_grid(region.era ~ species, scales="free") + xlab("SST (anomaly)")

dev.off()
```