---
title: "Nonstationary SST-salmon relationships"
author: "Mike Litzow"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
library(MuMIn)
library(dplyr)
library(ggplot2)
library(nlme)

options(max.print = 99999)

```

#Approach
Here I used mixed-effects models to test for changing SST effects by era. I fit separate models to each species, combining data over all three regions. I'm using the approach outlined by Zuur et al. 2009:

First, fit model with full fixed effects structure, using REML estimation, and select best random effects structure using AICc.
Second, refit with ML and compare nested fixed effects models to evaluate support for different fixed terms. This comparison uses the likelihood ratio test.
Third, refit selected model with REML to get unbiased coefficient estimates.

The full fixed structure that I'm using is:
ln(R/S) ~ intercept + stock:spawners + era + region + sst + sst:region:era 

Note that this model invokes a nonstationary sst effect for each region separately. I then test the nonstationary hypothesis by comparing with this reduced model:
ln(R/S) ~ intercept + stock:spawners + era + region + sst + sst:region

The reduced model invokes a stationary sst effect for each region.

I also explore improved fixed structures by iteratively dropping poorly-supported fixed effects. And I'm using lme with autocorrelated residuals throughout.

I am just making simple plots of each term at this point, rather than plotting a combined estimate of the different terms for each region/era combination.

Region = EBS/GOA/South, sst = sst at ocean entry, era = entry year before/after 1988/89

#Pink salmon
Begin by selecting the random effects structure.
```{r}
run.dat <- read.csv("coastwide data 1.17.19.csv")

# fixed effects model
p1 <- gls(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept for stocks
p2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept and slope on sst for stocks
p3 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# random intercept and era effect for stocks
p4 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and slope on sst by era for stocks
p5 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and region effect for stocks
p6 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

p7 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst for stocks
p8 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~ -1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst by era for stocks
p9 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random region effect for stocks
p10 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

p11 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

AIC <- AICc(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

So either p2 (random intercepts for each stock) or p4 (random intercepts and random era effects) is ok. Now look at random effects residuals for each.

```{r, echo=F}
qqnorm(p2, ~ranef(., level=1))
```

```{r, echo=F}
qqnorm(p4, ~ranef(., level=1))
```

Looks like they're about the same...I'll use p2 (random intercepts), as this is slightly favored.

Now test for the best-supported fixed structure using likelihood ratio test comparing nested models fit with ML. We'll compare a full model with a model replacing the sst:region:era interaction (invoking non-stationary sst effects in each region) with sst:region (invoking stationary sst effects in each region).

```{r}
# full model is p2 from above - but changing to ML estimation
pfull <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

p1 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

anova(pfull, p1)
```

So the full model, invoking nonstationary sst effects, is supported. I'll look at the t table for this supported model. Throughout, I'll be skipping the spawners:stock effects when printing results.

```{r, echo=F}
summary(pfull)$tTable[c(1:5,41:43),]
```

So, looking at the main terms, region and sst are supported, but a main era term is not. So we can see if a model excluding that term (but retaining the sst:region:era interaction) is better.

```{r}
p2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + sst:region:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

anova(pfull, p2)
```

Slightly better in AIC terms, but not supported by the likelihood ratio test. 

Now I'll fit the full model with REML and plot. 

```{r, echo=F}
pbest <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Pink"))

# summary(pbest)$tTable[c(1:5, 41:43),]

pplot <- as.data.frame(intervals(pbest, which="fixed")$fixed[c(2:5,41:43),])

pplot$term <- rownames(pplot)

pplot$term <- reorder(pplot$term, pplot$est.)

ggplot(pplot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Pink salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))

```

#Sockeye salmon
Again, begin with choosing the random structure.
```{r}

# fixed effects model
s1 <- gls(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# random intercept for stocks
s2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"))

# random intercept and slope on sst for stocks
s3 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and era effect for stocks
s4 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and slope on sst by era for stocks
s5 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and region effect for stocks
s6 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

s7 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst for stocks
s8 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~ -1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst by era for stocks
s9 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random region effect for stocks
s10 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

s11 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

AIC <- AICc(s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

So for sockeye a random intercept and random era factors is the best structure. Check the residuals.

```{r, echo=F}
qqnorm(s4, ~ranef(., level=1))
```

Pretty good? Check s2 (random intercept model) to compare.

```{r, echo=F}
qqnorm(s2, ~ranef(., level=1))
```

I don't think s2 is any better, and s4 is much better in terms of AICc, so I'll stick with s4.
```{r}
# full model is p2 from above - but changing to ML estimation
sfull <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + era | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

s1 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 + era | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

anova(sfull, s1)
```

So the reduced model, invoking stationary sst effects, is  supported. I'll refit with REML and look at the t table for this supported model.
```{r, echo=F}
s1 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 + era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

summary(s1)$tTable[c(1:5,37,36),]
```


 I'll try dropping the era term and see if that improves the model

```{r}
s1 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 + era | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

s2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + sst:region, random = ~1 + era | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

anova(s1, s2)

```

So s1, including the era term, is better. (And I realize we should have a fixed era term since we have a random era term?) I'll plot the best model (s1). 

```{r, echo=F}
sbest <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 + era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Sockeye"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))


splot <- as.data.frame(intervals(sbest, which="fixed")$fixed[c(2:5,36:37),])

splot$term <- rownames(splot)

splot$term <- reorder(splot$term, splot$est.)

ggplot(splot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Sockeye salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))

```

#Chum salmon
Again, begin with the random effects.
```{r}

# fixed effects model
c1 <- gls(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, method = "REML", 
          correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# random intercept for stocks
c2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# random intercept and slope on sst for stocks
c3 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and era effect for stocks
c4 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and slope on sst by era for stocks
c5 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random intercept and region effect for stocks
c6 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

c7 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst for stocks
c8 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~ -1 + sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random slope on sst by era for stocks
c9 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + sst:era | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

# random region effect for stocks
c10 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

c11 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~-1 + region:sst | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"),
          control=lmeControl(opt="optim", maxIter = 200, msMaxIter = 200))

AIC <- AICc(c1, c2, c3, c4, c5, c6, c7, c8, c9, c10, c11) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

c2 (random intercept model) is slightly better than c4 (random intercepts and era terms). Look at the residuals for each.

```{r, echo=F}
qqnorm(c2, ~ranef(., level=1))
```

That looks fine. So we'll examine c4 just for comparison.

```{r, echo=F}
qqnorm(c4, ~ranef(., level=1))
```

That's worse. So c2 it is.

Compare nested fixed structure to test hypothesis of nonstationary sst effects.

```{r}
# full model is p2 from above - but changing to ML estimation
cfull <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

c1 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c1)
```

So here, again, the nonstationary model is best. Here are the t-table results.

```{r, echo=F}
cfull <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

summary(cfull)$tTable[c(1:5, 27:29),]
```

So the era and region main effects are both poorly supported. We'll see if dropping these is supported.

```{r}
# refit full model with ML estimation
cfull <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

# model with sst:era effect across all regions
c2 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c2)
```

No support. I'll see if a model without the region term is supported

```{r}
# model with sst:era effect across all regions
c3 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + era + sst:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c3)
```

Full model has the better AICc score (although the BIC score is lower for the reduced model!). Try a model without a era main effect.

```{r}
# model with sst:era effect across all regions
c4 <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + sst:era, random = ~1 | stock,
          method = "ML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

anova(cfull, c4)
```

Still no improvement in AICc terms. So we'll accept the full model for now. Refit with REML and plot.

```{r, echo=F}
cbest <- lme(log(recruits/spawners) ~ 1 + stock:spawners + sst + region + era + sst:region:era, random = ~1 | stock,
          method = "REML", correlation=corAR1(form= ~ 1 | stock), data = filter(run.dat, species=="Chum"))

cplot <- as.data.frame(intervals(cbest, which="fixed")$fixed[c(2:5,27:29),])

cplot$term <- rownames(cplot)

cplot$term <- reorder(cplot$term, cplot$est.)

ggplot(cplot, aes(term, est.)) + geom_errorbar(aes(ymax=upper, ymin=lower), width=0.2) + geom_point(size=3) + coord_flip() +
  xlab("") + ylab("Coefficient") +  ggtitle("Chum salmon") + geom_hline(yintercept = 0) + 
  theme(axis.text.y  = element_text(size=10))

```

No obvious difference in the regional effect of sst in the different eras, which makes it seem that c3 would have been a better model. I might be making a mistake in interpreting the likelihood ratio results?